ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-02 rebuttal.tex

    The line: `if the ChemKED file encodes an RCM experiment' is unnecessarily repeated.
    %
\end{reviewer}

\begin{response}
    %
    We have consolidated the RCM-specific information into a separate list, and eliminated the
    duplicated phrase.
    %
\end{response}

\begin{reviewer}
    %
    21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online
    lookup''. There are two main caveats with online lookups: a) It will make the code slower if the
    internet connection is not strong, b) an internet connection is necessary for the code to be
    executed.
    %
\end{reviewer}

\begin{response}
    %
    In general, both of these points are correct. However, we believe there are several mitigating
    factors that make the online lookups worthwhile. First, we explicitly test that our code fails
    gracefully in the absence of an internet connection. No internet connection is necessary to run
    the code at this time, and aside from specialized features, we do not anticipate that an
    internet connection will ever be required to run the basic software. Second, for the specific
    feature mentioned here, this will be something that the user is explicitly requesting (e.g., by
    attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for
    the connection to complete to use the feature. Third, we expect that the amount of data
    transmitted for such a request would be small, on the order of a few kB, so that even slow
    internet connections would be able to handle such requests. Finally, in addition to
    internet-based options, we are also investigating the use of standalone packages, such as RDKit,
    that would enable local retrieval of such information. We have not modified the text in response
    to the comment.
    %
\end{response}

### 2017-10-02 rebuttal.tex

\begin{reviewer}
    %
    21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online
    lookup''. There are two main caveats with online lookups: a) It will make the code slower if the
    internet connection is not strong, b) an internet connection is necessary for the code to be
    executed.
    %
\end{reviewer}

\begin{response}
    %
    In general, both of these points are correct. However, we believe there are several mitigating
    factors that make the online lookups worthwhile. First, we explicitly test that our code fails
    gracefully in the absence of an internet connection. No internet connection is necessary to run
    the code at this time, and aside from specialized features, we do not anticipate that an
    internet connection will ever be required to run the basic software. Second, for the specific
    feature mentioned here, this will be something that the user is explicitly requesting (e.g., by
    attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for
    the connection to complete to use the feature. Third, we expect that the amount of data
    transmitted for such a request would be small, on the order of a few kB, so that even slow
    internet connections would be able to handle such requests. Finally, in addition to
    internet-based options, we are also investigating the use of standalone packages, such as RDKit,
    that would enable local retrieval of such information. We have not modified the text in response
    to the comment.
    %
\end{response}

\begin{reviewer}
    %
    22.~Page 7, line 40: Why is it interesting to create the YAML file again with the
    \texttt{write\_file()} function. Creating a summary excel/CSV file would be more useful.
    %
\end{reviewer}

\begin{response}
    %
    The use case for the \texttt{write\_file()} is that users might modify the ChemKED instance
    after it has been loaded, and wish to write their changes out to a new file.

    Users can create Excel or CSV files containing all or some of the datapoints in the ChemKED
    instance by using the \verb|get_dataframe()| function, and the associated functionality in
    Pandas (\verb|DataFrame.to_csv()| and \verb|DataFrame.to_excel()|). Since this functionality
    already exists, we do not wish to reinvent the wheel by writing our own functions that would
    have to be maintained and tested. Finally, we would note that data can also be plotted and
    viewed directly from Python using existing functionality in the Pandas DataFrame class
    (\verb|DataFrame.plot()|).
    %
\end{response}

### 2017-10-02 pyked.tex
WIP respond to #39
Modify the text to remove "might"

The \ck{} file that will be used in this example can be found in the
\texttt{tests} directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: 0
chemked-version: 0.1.6
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, this file contains just a single datapoint, which describes the experimental
ignition delay, initial mixture composition, initial temperature, initial pressure,
compression time, ignition type, and volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
- temperature:
    - 297.4 kelvin
  ignition-delay:
    - 1.0 ms
  pressure:
    - 958.0 torr
  composition:
    kind: mole fraction
    species:
      - species-name: H2
        InChI: 1S/H2/h1H
        amount:
          - 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        amount:
          - 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        amount:
          - 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        amount:
          - 0.63125
  ignition-type:
    target: pressure
    type: d/dt max
  compression-time:
    - 38.0 ms
  volume-history:
    time:
      units: s
      column: 0
    volume:
      units: cm3
      column: 1
    values:
      - [0.00E+000, 5.47669375000E+002]
      - [1.00E-003, 5.46608789894E+002]
      - [2.00E-003, 5.43427034574E+002]
      ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. A user
might want to load the information from this file to perform a simulation using
Cantera~\autocite{Cantera:2.3.0} to calculate the ignition delay. All information
required to perform this simulation is present in the \ck{} file, with the exception
of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

>>>>

The \ck{} file that will be used in this example can be found in the
\texttt{tests} directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: 0
chemked-version: 0.1.6
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%r
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, this file contains just a single datapoint, which describes the experimental
ignition delay, initial mixture composition, initial temperature, initial pressure,
compression time, ignition type, and volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
- temperature:
    - 297.4 kelvin
  ignition-delay:
    - 1.0 ms
  pressure:
    - 958.0 torr
  composition:
    kind: mole fraction
    species:
      - species-name: H2
        InChI: 1S/H2/h1H
        amount:
          - 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        amount:
          - 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        amount:
          - 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        amount:
          - 0.63125
  ignition-type:
    target: pressure
    type: d/dt max
  compression-time:
    - 38.0 ms
  volume-history:
    time:
      units: s
      column: 0
    volume:
      units: cm3
      column: 1
    values:
      - [0.00E+000, 5.47669375000E+002]
      - [1.00E-003, 5.46608789894E+002]
      - [2.00E-003, 5.43427034574E+002]
      ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. One application of the
data stored in this file is to perform a simulation using Cantera~\autocite{Cantera:2.3.0} to
calculate the ignition delay, including the facility-dependent effects represented in the volume
trace. All information required to perform this simulation is present in the \ck{} file, with the
exception of a chemical kinetic model for \ce{H2}\slash {CO} combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED
